DRAFT_Cleaning

### 2024-12-09 Draft_Cleaning.Rmd
copied existing data over

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(caret)
library(ggformula)

@@

gage_data2<- gage_data |>
  group_by(Date)|> 
   summarize(across(everything(),~first(na.omit(.)), .names = "{.col}"))|>
  mutate(DATE_START= as_date(dmy(Date)))|>
    arrange(DATE_START) 
gage_data2<-gage_data2|>
    mutate(across(c(-Date,-DATE_START), 
                ~ ifelse(is.na(.), 
                         coalesce((lag(.) + lead(.)) / 2, lag(.), lead(.)), 
                         .)))|>
  select(-Date)

@@

dredge_PCA = prcomp(dredge_prcomp, scale = T)
biplot(dredge_PCA)

@@

###xgBoost
```{r}
library(xgboost)
gage_xgb <- gage_survey |>
  select(-Date_obvs,-Date_7,-Date_14,-DATE_START,
         -lead_7,-lead_14,-SURVEYJOBIDPK,-EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 10) # Increase this to 10 to get more consistent RMSEs
fit_dredge = train(VOLUMEDREDGED ~ ., 
                 data = gage_xgb,
                 method = "xgbTree",
                 tuneGrid = expand.grid(nrounds = c (15,25,50,100), max_depth = 1:4, 
                                        eta = 0.3, gamma = 0, 
                                        colsample_bytree = .8, 
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0, 
                 trControl = ctrl, 
                 na.action = na.exclude)

@@

@@

  ctrl = trainControl(method = "cv", number = 5)
  fit_dredge_DCV = train(VOLUMEDREDGED ~ ., 
                       data = gage_xgb[in_train, ],
                       method = "xgbTree",
                       tuneGrid = expand.grid(nrounds = c (15,25,50,100), max_depth = 1:6,
                                        eta = 0.3, gamma = 0, 
                                        colsample_bytree = 1, 
                                        min_child_weight = 1, subsample = 1),
                       verbosity = 0, 
                       trControl = ctrl,
                       na.action = na.exclude)

@@

fit_dredge_DCV

confusion_matrix <- confusionMatrix(cv_pred[in_test], gage_xgb[in_test, ]$VOLUMEDREDGED)
print(confusion_matrix)

>>>>

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(caret)
library(ggfortify)
library(ggformula)

@@

gage_data2<- gage_data |>
  group_by(Date)|>
   summarize(across(everything(),~first(na.omit(.)), .names = "{.col}"))|>
  mutate(DATE_START= as_date(dmy(Date)))|>
    arrange(DATE_START)

gage_data2<-gage_data2|>
    mutate(across(c(-Date,-DATE_START),
                ~ ifelse(is.na(.),
                         coalesce((lag(.) + lead(.)) / 2, lag(.), lead(.)),
                         .)))|>
  select(-Date)

@@

dredge_PCA = prcomp(dredge_prcomp, scale = T)

biplot(dredge_PCA)

@@

###xgBoost
```{r}
library(xgboost)
gage_xgb <- gage_survey |>
  select(-lead_7, -lead_14,-DATE_START,
         -SURVEYJOBIDPK,-EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 10) 
fit_dredge = train(VOLUMEDREDGED ~ .,
                 data = gage_xgb,
                 method = "xgbTree",
                 tuneGrid = expand.grid(nrounds = c (15,25,50,100), max_depth = 1:10,
                                        eta = 0.3, gamma = 0,
                                        colsample_bytree = .8,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

@@

fit_dredge

@@

  ctrl = trainControl(method = "cv", number = 10)
  fit_dredge_DCV = train(VOLUMEDREDGED ~ .,
                       data = gage_xgb[in_train, ],
                       method = "xgbTree",
                       tuneGrid = expand.grid(nrounds = c (15,25,50,100), max_depth = 1:10,
                                        eta = 0.3, gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                       verbosity = 0,
                       trControl = ctrl,
                       na.action = na.exclude)

@@

tune_grid <- expand.grid(
  nrounds = c(50, 100, 150), 
  max_depth = c(3, 6, 9), 
  eta = c(0.1, 0.2, 0.3), 
  gamma = c(0, 0.1), 
  colsample_bytree = c(0.6, 0.8), 
  min_child_weight = c(1, 2, 5), 
  subsample = c(0.8, 1)
)

set.seed(123)
ctrl = trainControl(method = "cv", number = 10, savePredictions = "final")

fit_dredge2 = train(VOLUMEDREDGED ~ .,
                   data = gage_xgb,
                   method = "xgbTree",
                   tuneGrid = tune_grid,
                   verbosity = 0,
                   trControl = ctrl,
                   na.action = na.exclude)

# Model performance
fit_dredge2

# Model evaluation
predictions <- fit_dredge2$pred  # Predictions from cross-validation
head(predictions)

# RMSE, R-Squared, etc.
cv_rmse <- sqrt(mean((predictions$obs - predictions$pred)^2))
cv_r2 <- cor(predictions$obs, predictions$pred)^2

cat("RMSE: ", cv_rmse, "\n")
cat("R-Squared: ", cv_r2, "\n")

##Og
train<-gage_xgb[in_train, ]
best_tune<-fit_dredge_DCV$results
best_tune
confusion_matrix <- confusionMatrix(cv_pred[in_test], gage_xgb[in_test, ])
print(confusion_matrix)
